---
title: "R Notebook"
author: "Shahen Yashpal"
output: 
    html_document:
        toc: true
        toc_float:
            collapsed: false
        theme: yeti
        highlight: haddock
---
# Git Basic Commands

- `git status` in git this is the smae thing as telling me the status
- `git add` equivalent to checking the box
- `git commit` this commed with -m "" message to be added when commiting
- `git log` shows history
- `git diff` shows difference since last commit

## Secondary commands

Infrequently Used commands

- `git remote`

More Commonly Used:

- `git push`
- `git pull`

# Basic R Commands

## Basic Math

```{r}
1 + 1
1 ^ 1
```
## Data Wrangler

```{r}
library(tidyverse)
```

- `select`
- `arrange`
- `mutate`
- `group_by`
- `summarize`
- `filter`
- `rename`
- ` %>% `

```{r}
class(1.23)
```

```{r}
class(1:10)
```

```{r}
CO2_df <- tbl_df(CO2)
class(CO2_df)
```
```{r}
CO2_df
```

```{r}
CO2_df %>% 
    select(Plant, Type, conc)
```
```{r}
names(CO2_df)
```
```{r}
CO2_df %>% select(Plant, Type)
```

```{r}
glimpse (CO2_df) #tells me what my data looks like
str(CO2_df) #tells me in more detail what my data looks like
str(1:10)
```


```{r}
CO2_df %>%  arrange(conc)
CO2_df %>%  arrange(desc(conc))
CO2_df %>%  arrange(Type, desc(conc))
```

