---
title: "R Notebook"
author: "Luke Johnston"
output: html_notebook
---

# Git basic commands

- `git status`: 
- `git add`
- `git commit`
- `git log`
- `git diff`

## Secondary commands

Infrequently used:

- `git remote`

More commonly used:

- `git push`
- `git pull`

# Basic R commands

## Basic math

```{r}
1 + 1
1 - 1
1 * 1
1 / 1
1 ^ 1
```

## Data wrangling with dplyr

### dplyr

```{r}
library(tidyverse)
```

- `select`
- `arrange`
- `mutate`
- `group_by`
- `summarize`
- `filter`
- `rename`
- `%>%`
- `tbl_df`

```{r}
class(CO2)
```

```{r}
class(mean)
```

```{r}
class(1:10)
```

```{r}
class(1.23)
```

```{r}
CO2_df <- tbl_df(CO2)
class(CO2_df)
```

```{r}
CO2
CO2_df
```

```{r}
names(CO2_df)
```


```{r}
CO2_df %>%
    select(Plant, Type, conc)
```

```{r}
CO2_df %>% select(Plant, Type) %>% select(Plant)
select(select(CO2_df, Plant, Type), Plant)
CO2_df %>% select(., Plant)
```

```{r}
str(CO2_df)
```

```{r}
glimpse(CO2_df)
```

```{r}
CO2_df %>% arrange(conc)
CO2_df %>% arrange(desc(conc))
CO2_df %>% arrange(Type, desc(conc))
```

```{r}
CO2_df %>%
    mutate(Test = "Yes",
           conc_small = conc / 100,
           Treatment_sp = ifelse(Treatment == "nonchilled", "non-chilled", Treatment),
           lconc_small = log(conc_small))

# mutate_if mutate_at mutate_all mutate_each
```

```{r}
CO2_df %>% 
    rename(plant = Plant, type = Type)
```

```{r}
TRUE
FALSE
class(TRUE)

TRUE & TRUE
TRUE | TRUE
TRUE & FALSE
TRUE | FALSE
FALSE | FALSE
FALSE & FALSE
TRUE | (FALSE & TRUE)
(TRUE | TRUE) & FALSE
```


```{r}
CO2_df %>% 
    filter(Type == "Quebec")

CO2_df[CO2_df$Type == "Quebec", ]
```

```{r}
CO2_df %>% 
    filter(Type == "Quebec" & Treatment == "chilled")

CO2_df %>% 
    filter(Type == "Quebec" | Treatment == "chilled")

CO2_df %>% 
    filter(Plant == "Qn1" | Plant == "Qn2")

CO2_df %>% 
    filter(Plant == "Qn1" & Plant == "Qn2")

CO2_df %>% 
    filter(Plant %in% c("Qn1", "Qn2", "Qn3"))

CO2_df %>% 
    filter(conc > 300)

# > < >= <= ==

CO2_df %>% 
    filter(between(conc, 300, 700))
```

hello
```{r}
names(iris)
```

```{r}
library(tidyverse)
iris %>% 
    filter(Petal.Length > 2.5) %>% 
    group_by(Species) %>% 
    summarize(mean.sepal.length = mean(Sepal.Length), 
              sd.sepal.length = sd(Sepal.Length),
              mean.sepal.width = mean(Sepal.Width),
              sd.sepal.width = mean(Sepal.Width))

iris %>% 
    filter(Petal.Length > 2.5) %>% 
    group_by(Species) %>% 
    summarise_at(vars(starts_with("Sepal")),funs(mean, sd)) #faster way of doing the above code which was like 4 lines
        
```

### tidyr

-gather #takes Wide and make sit into Long
- Spread #takes Long and changes it into Wide

```{r}
ds <- data.frame(
x1 = 1:10,
x2 = 11:20,
x3 = 21:30
)
ds %>% gather(Variable, Values)

```

