---
title: "R Notebook"
author: "Luke Johnston"
output: html_notebook
---

# Git basic commands

- `git status`: 
- `git add`
- `git commit`
- `git log`
- `git diff`

## Secondary commands

Infrequently used:

- `git remote`

More commonly used:

- `git push`
- `git pull`

# Basic R commands

## Basic math

```{r}
1 + 1
1 - 1
1 * 1
1 / 1
1 ^ 1
```

## Data wrangling

### With dplyr

### dplyr

```{r}
library(tidyverse)
```

- `select`
- `arrange`
- `mutate`
- `group_by`
- `summarize`
- `filter`
- `rename`
- `%>%`
- `tbl_df`

```{r}
class(CO2)
```

```{r}
class(mean)
```

```{r}
class(1:10)
```

```{r}
class(1.23)
```

```{r}
CO2_df <- tbl_df(CO2)
class(CO2_df)
```

```{r}
CO2
CO2_df
```

```{r}
names(CO2_df)
```


```{r}
CO2_df %>%
    select(Plant, Type, conc)
```

```{r}
CO2_df %>% select(Plant, Type) %>% select(Plant)
select(select(CO2_df, Plant, Type), Plant)
CO2_df %>% select(., Plant)
```

```{r}
str(CO2_df)
```

```{r}
glimpse(CO2_df)
```

```{r}
CO2_df %>% arrange(conc)
CO2_df %>% arrange(desc(conc))
CO2_df %>% arrange(Type, desc(conc))
```

```{r}
CO2_df %>%
    mutate(Test = "Yes",
           conc_small = conc / 100,
           Treatment_sp = ifelse(Treatment == "nonchilled", "non-chilled", Treatment),
           lconc_small = log(conc_small))

# mutate_if mutate_at mutate_all mutate_each
```

```{r}
CO2_df %>% 
    rename(plant = Plant, type = Type)
```

```{r}
TRUE
FALSE
class(TRUE)

TRUE & TRUE
TRUE | TRUE
TRUE & FALSE
TRUE | FALSE
FALSE | FALSE
FALSE & FALSE
TRUE | (FALSE & TRUE)
(TRUE | TRUE) & FALSE
```


```{r}
CO2_df %>% 
    filter(Type == "Quebec")

CO2_df[CO2_df$Type == "Quebec", ]
```

```{r}
CO2_df %>% 
    filter(Type == "Quebec" & Treatment == "chilled")

CO2_df %>% 
    filter(Type == "Quebec" | Treatment == "chilled")

CO2_df %>% 
    filter(Plant == "Qn1" | Plant == "Qn2")

CO2_df %>% 
    filter(Plant == "Qn1" & Plant == "Qn2")

CO2_df %>% 
    filter(Plant %in% c("Qn1", "Qn2", "Qn3"))

CO2_df %>% 
    filter(conc > 300)

# > < >= <= ==

CO2_df %>% 
    filter(between(conc, 300, 700))
```

```{r}
library(tidyverse)
co2df <- tbl_df(CO2)

names(co2df)
co2df %>% 
    summarize(mean_conc = mean(conc),
              mean_uptake = mean(uptake))
mean(co2df$conc)

co2df %>% 
    group_by(Type) %>% 
    summarize(mean_conc = mean(conc),
              mean_uptake = mean(uptake))

mean(co2df[co2df$Type == "Quebec", ]$uptake)
mean(co2df[co2df$Type != "Quebec", ]$uptake)
    
co2df %>% 
    group_by(Type, Treatment) %>% 
    summarize(mean_uptake = sum(uptake))
```

Use iris dataset, keep only `Petal.Length` > 2.5, and find the `mean` and `sd`
of `Sepal.Length` and `Sepal.Width` by `Species`, the tidyverse way.


```{r}
iris %>% 
    filter(Petal.Length > 2.5) %>% 
    group_by(Species) %>% 
    summarise_all(funs(mean, sd))

```

### tidyr

- `gather`: Wide -> Long
- `spread`: Long -> Wide

```{r}
ds <- data.frame(
    x1 = 1:10,
    x2 = 11:20,
    x3 = 21:30
) 
ds
ds %>% 
    gather(Variable, Values) 
# %>% 
    # group_by(Variable) %>% 
    # summarize_all(funs(mean))

```


```{r}
mean_sd <- function(x, digits = 2) {
    mean_val <- round(mean(x), digits)
    sd_val <- round(sd(x), digits)
    paste0(mean_val, " (", sd_val, ")")
}

iris %>% 
    gather(Variable, Values, -Species) %>% 
    group_by(Species, Variable) %>% 
    summarize(MeanSD = mean_sd(Values, 4)) %>% 
    spread(Species, MeanSD)
    
```

## Visualization

```{r}
library(tidyverse)

names(diamonds)
```

```{r}
glimpse(diamonds)
```


```{r}
dim(diamonds)
```

```{r}
ggplot(diamonds)
```

```{r}
ggplot(diamonds, aes(x = carat, y = price))
```

```{r}
ggplot(diamonds, aes(x = carat, y = price)) +
    geom_point()
```

```{r}
ggplot(diamonds, aes(x = carat, y = price, color = cut)) +
    geom_point(alpha = 0.3) +
    facet_wrap(~ clarity, ncol = 3)
```

```{r}
names(swiss)
```

```{r}
ggplot(swiss, aes(x = Fertility)) +
    geom_density()
```

```{r}
swiss %>% 
    gather(Variable, Value) %>% 
    ggplot(aes(x = Value)) +
    geom_density() +
    facet_wrap(~ Variable)
```

```{r}
swiss %>% 
    gather(Variable, Value) %>% 
    ggplot(aes(x = Variable, y = Value)) +
    geom_boxplot()
```

```{r}
swiss %>% 
    gather(Variable, Value) %>% 
    ggplot(aes(x = Variable, y = Value, fill = Variable)) +
    geom_boxplot() +
    scale_fill_manual(name = "Dataset columns", 
                      values = c("red", "green", "blue", "black", "yellow", "pink")) +
    theme(axis.line = element_line(color = "black"),
          panel.background = element_blank(),
          axis.text.x = element_text(size = 5)) +
    labs(x = "Variables within dataset", y = "Number", title = "This is an example")
```

```{r}
swiss %>% 
    gather(Variable, Value) %>% 
    ggplot(aes(x = Value)) +
    geom_density() +
    facet_wrap(~ Variable, scales = "free", strip.position = "right")
```

